Elephant Seal Age JAE Intraspecific Variation

### 2023-09-20 paper/paper.qmd
Made the annual timing figure.

@@

As maternal age increases, pup sex skews slightly towards males. Pup survival declines slightly with maternal age, but pup recruitment is not significantly affected by maternal age (@fig-repro_age). Reproductive declines therefore do not persist to the next generation.

@@

In order to create the timing figure, I need a csv with these columns: breeding arrival DOY, breeding departure DOY, molt arrival DOY, molt departure DOY.

-   breeding arrival DOY = firstobsbreeddoy

-   breeding departure DOY = firstobsbreeddoy + breeddur

-   molt arrival DOY = lastobsmoltdoy - moltdur

-   molt departure DOY = lastobsmoltdoy

```{r}
#| label: annual_timing

@@

##

>>>>

We predicted that breeding success would decline with age. We found that breeding success increases to prime age, then decreases at older ages (@fig-age_dist).

@@

We predicted that reproductive declines would carry over to the next generation.

We found that as maternal age increases, pup sex skews slightly towards males. Pup survival declines slightly with maternal age, and pup recruitment is not significantly affected by maternal age (@fig-repro_age).

We reject our hypothesis: reproductive declines do not persist to the next generation.

@@

We predicted that seals would optimize the timing of their annual cycle.

We found that as maternal age increases, there are slight adjustments to the timing of life history events. These events (breeding/molting arrival/departure) are typically shifted later for older animals (@fig-annual_timing).

```{r}
#| label: fig-annual_timing
#| fig-cap: Timing of life history events across maternal age classes.

sealdat %>% 
  mutate(ba = firstobsbreeddoy, 
         bd = firstobsbreeddoy + breeddur, 
         ma = lastobsmoltdoy - moltdur, 
         md = lastobsmoltdoy) %>% 
  drop_na(age, ba, bd, ma, md) %>% 
  group_by(age) %>% 
  summarize(across(ba:md, mean)) %>% 
  pivot_longer(cols = -age, 
               names_to = "event", 
               values_to = "doy") %>% 
  group_by(age) %>% 
  group_modify(\(data, keys) {
    tibble(event = c("Long trip", "Breeding", "Short trip", "Molting", "longtrip2"),
           doy = c(data$doy, 365) - c(0, data$doy))
  }) %>% 
  ungroup() %>% 
  mutate(event = factor(event, levels = c("Long trip", 
                                          "Breeding", 
                                          "Short trip", 
                                          "Molting", 
                                          "longtrip2"))) %>% 
  ggplot(aes(x = age, y = doy, fill = event)) + 
  geom_area(position = position_stack(reverse = TRUE),
            outline.type = "full",
            color = "white") +
  scale_fill_discrete(type = c("#829BD1", "#E9B668", "#7EC6CC", "#E091B9", "#829BD1"),
                      breaks = c("Long trip", "Breeding", "Short trip", "Molting")) + 
  labs(x = "Maternal Age (Years)", y = "Day of Year") + 
  theme_few() + 
  theme(legend.position = "right",
        legend.title = element_blank())

@@
